Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-03-20 paper/manuscript.Rmd
Update manuscript.Rmd and manuscript.md

In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Only `r n10` of these 12 individuals had more than 10
swimming events, suggesting that for ~10% of collared individuals swim
relatively frequently, on average occurring approximately every `r indMean` days
during ice-free seasons. We did not detect any swimming events for the remaining
17 individuals. It remains unclear whether these individuals do not swim, or
whether swimming is rare, and thus undetected, for the majority of individuals
Overall, there appears to be three groups of individuals: those that swim often
(n = 3, every ~`r indMean` days), those that swim occasionally (n = 9, every ~`r
ind2Mean` days), and those that never swam (n = 17). We are also unable to draw
conclusions about the prevalence of swimming in the ocean by male caribou: our
initial visual observation was of an adult male, but our GPS data only included
females.

@@

Oceanic swimming is relatively common among a minority of
marked caribou. Of the three collared individuals that swam the most, two were collared on
Western Indian Island and the third was collared on Fogo Island. Inter-island
distances between Western and Eastern Indian Islands, and the smaller islands
around them, are shorter than those between Fogo Island and others (Figure 2). It is possible that caribou display natal philopatry [@Larue2019] to some of the smaller islands and thus are more likely to  
swim between them more frequently than animals that were born, ospend most of their time, on
Fogo Island. The forage limitation hypothesis is generally thought to apply
uniformly across populations, but individual caribou could evaluate the costs
and benefits of swimming to another island differently. It is possible that some
individuals, in our case ~10%, considered swimming to be less costly or
perceived competition or predation more acutely than their conspecifics.

For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and it is unknown when caribou began to occupy other
islands, anecdotal evidence suggests it was during the 1990s. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation, our data suggest caribou only remain on smaller islands for a few
days at a time, so the fitness equilibrium remains a theoretical construct. More
likely, however, is that fine-scale competition, density-dependent habitat
selection, and forage depletion drives individuals to peridiodically swim
between islands. Given the presumed low density of coyotes on Fogo Island and
the minimal risk they pose, predation risk does not appear to be a major driver
of swimming.

>>>>

In total, 12 of 29 collared female caribou swam between islands in the Fogo
Island archipelago. Only `r n10` of these 12 individuals had more than 10
swimming events, suggesting that for ~10% of collared individuals swim
relatively frequently, on average occurring approximately every `r indMean` days
during ice-free seasons. We did not detect any swimming events for the remaining
17 individuals. It remains unclear whether these individuals do not swim, or
whether swimming is rare, and thus undetected, for the majority of individuals
Overall, there appears to be three groups of individuals: those that swim often
(n = 3, every ~`r indMean` days), those that swim occasionally (n = 9, every ~
`r ind2Mean` days), and those that never swam (n = 17). We are also unable to draw
conclusions about the prevalence of swimming in the ocean by male caribou: our
initial visual observation was of an adult male, but our GPS data only included
females.

@@

Oceanic swimming is relatively common among a minority of marked caribou. Of the
three collared individuals that swam the most, two were collared on Western
Indian Island and the third was collared on Fogo Island. Inter-island distances
between Western and Eastern Indian Islands, and the smaller islands around them,
are shorter than those between Fogo Island and others (Figure 2). It is possible
that caribou display natal philopatry [e.g. @Larue2018] to some of the smaller
islands and thus are more likely to swim between them more frequently than
animals that were born, ospend most of their time, on Fogo Island. The forage
limitation hypothesis is generally thought to apply uniformly across
populations, but individual caribou could evaluate the costs and benefits of
swimming to another island differently. It is possible that some individuals, in
our case ~10%, considered swimming to be less costly or perceived competition or
predation more acutely than their conspecifics.

For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and it is unknown when caribou began to occupy other
islands, although anecdotal evidence suggests it was during the 1990s. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation, our data suggest caribou only remain on smaller islands for a few
days at a time, so the fitness equilibrium remains a theoretical construct. More
likely, however, is that fine-scale competition, density-dependent habitat
selection, and forage depletion drives individuals to peridiodically swim
between islands. Given the presumed low density of coyotes on Fogo Island and
the minimal risk they pose, predation risk does not appear to be a major driver
of swimming.
